Lesson 6
========================================================

### Welcome
Notes:Here we attempt to deduce the factors that go into the price of a diamonds. Socio-Economic and Political history of diamonds industry is fascinating. Diamonds gave rise to mining industry in south Africa. Diamonds were the reason of the colonization of South Africa and both historical and mordern wars and unrest in the region.

***

### Scatterplot Review

```{r Scatterplot Review}
library(ggplot2)
data(diamonds)
names(diamonds)
ggplot(data=diamonds,aes(y=price,x=carat))+
  geom_point(alpha=1/10,aes(color='black',fill='red'))+
  coord_cartesian(xlim=c(0,quantile(diamonds$carat,0.99)),
                  ylim=c(0,quantile(diamonds$price,0.99)))
ggsave('1_price_vs_carat.png')
```

***

### Price and Carat Relationship
Response:
price vs carat is having a positive linear relationship.
***

### Frances Gerety
Notes:

#### A diamonds is'forever'!!


***

### The Rise of Diamonds
Notes:

***

### ggpairs Function
Notes:

```{r ggpairs Function}
# install these if necessary
install.packages('GGally')
install.packages('scales')
install.packages('memisc')
install.packages('lattice')
install.packages('MASS')
install.packages('car')
install.packages('reshape')
install.packages('plyr')

# load the ggplot graphics package and the others
library(ggplot2)
library(GGally)
library(scales)
library(memisc)

# sample 10,000 diamonds from the data set
set.seed(20022012)
diamond_samp <- diamonds[sample(1:length(diamonds$price), 10000), ]
ggpairs(diamond_samp, 
        lower = list(continuous = wrap("points", shape = I('.'))), 
        upper = list(combo = wrap("box", outlier.shape = I('.'))))

ggpairs(diamond_samp)
ggsave('2_ggpairs.png')

```

What are some things you notice in the ggpairs output?
Response:
Carat is strongly related to the price of diamond and it's x, y and z measure. and price is also strongly correlated to the volume measures x,y,and z have almost a linear relationship There's a pattern in relation of table with price and volume measures. Check the relationships between (price and clarity) and (price and color).

Weight(xyz)~Price(exponentiality in the first plot, because on the demand side, the big chunks of diamonds with no significant flaws are harder to find and hence, higher the price.)
***

### The Demand of Diamonds
Notes:

```{r The Demand of Diamonds}
library(gridExtra)

plot1 <- qplot(data=diamonds,x=price,binwidth=100,color=I('Black'),fill=I('#099DD9')) + 
  ggtitle('Price')

plot2 <- qplot(data=diamonds,x=price,binwidth=0.01,color=I('Black'),fill=I('#F79420')) + scale_x_log10()+
  ggtitle('Price (log10)')

grid.arrange(plot1,plot2)
ggsave('3_price_histogram.png')
#customers on smaller side of the price and small size of diamonds are more sensitive to price than 
#other well to do customers. THey(less than one carat customers) won't buy the diamond if it werent #customary and there are fewer customers who can afford a bigger diamonds so market is not that #competitive interms of price as we increases the carat size,therefore the increase in the variance , #it may also be result of some other path dependence,multiplicative processes or some combination of #both.Hence it's good idea to use log scale
```

***

### Connecting Demand and Price Distributions
Notes:

***

### Scatterplot Transformation

```{r Scatterplot Transformation}
p1<-ggplot(data=diamonds,aes(y=price,x=carat))+geom_point(alpha=1/10,aes(color='black',fill='red'))+
  coord_cartesian(xlim=c(0,quantile(diamonds$carat,0.99)),
                  ylim=c(0,quantile(diamonds$price,0.99)))+
  ggtitle('carat vs price')
  
  
p2<-ggplot(data=diamonds,aes(y=price,x=carat))+geom_point(alpha=1/10,aes(color='black',fill='red'))+
  coord_cartesian(xlim=c(0,quantile(diamonds$carat,0.99)),
                  ylim=c(0,quantile(diamonds$price,0.99)))+
  ggtitle('carat vs log10(price)')+
  coord_trans(y='log10')

grid.arrange(p1,p2)
ggsave('4_price_carat_log10.png')
```


### Create a new function to transform the carat variable

```{r cuberoot transformation}
cuberoot_trans = function() trans_new('cuberoot', transform = function(x) x^(1/3),
                                      inverse = function(x) x^3)
```

#### Use the cuberoot_trans function
```{r Use cuberoot_trans}
ggplot(aes(carat, price), data = diamonds) + 
  geom_point() + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

***

### Overplotting Revisited

```{r Sort and Head Tables}
ggplot(aes(carat, price), data = diamonds) + 
  geom_point(size=3/4, alpha=1/2,position = position_jitter()) + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
ggsave('6_Price_(log10)_by_Cube-Root_of_Carat.png')
```


```{r Overplotting Revisited}
ggplot(aes(carat, price), data = diamonds) + 
  geom_point() + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

***

### Other Qualitative Factors
Notes:

***

### Price vs. Carat and Clarity

Alter the code below.
```{r Price vs. Carat and Clarity}
# install and load the RColorBrewer package
install.packages('RColorBrewer')
library(RColorBrewer)

ggplot(aes(x = carat, y = price), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
    guide = guide_legend(title = 'Clarity', reverse = T,
    override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
    breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
    breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')
```

***

### Clarity and Price
Response:

***

### Price vs. Carat and Cut

Alter the code below.
```{r Price vs. Carat and Cut}
ggplot(aes(x = carat, y = price, color = clarity), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Clarity', reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')
```

***

### Cut and Price
Response:

***

### Price vs. Carat and Color

Alter the code below.
```{r Price vs. Carat and Color}
ggplot(aes(x = carat, y = price, color = cut), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = Cut, reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Cut')
```

***

### Color and Price
Response:

***

### Linear Models in R
Notes:

Response:

***

### Building the Linear Model
Notes:

```{r Building the Linear Model}
m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = diamonds)
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1, m2, m3, m4, m5)
```

Notice how adding cut to our model does not help explain much of the variance
in the price of diamonds. This fits with out exploration earlier.

***

### Model Problems
Video Notes:

Research:
(Take some time to come up with 2-4 problems for the model)
(You should 10-20 min on this)

Response:

***

### A Bigger, Better Data Set
Notes:

```{r A Bigger, Better Data Set}
install.package('bitops')
install.packages('RCurl')
library('bitops')
library('RCurl')

diamondsurl = getBinaryURL("https://raw.github.com/solomonm/diamonds-data/master/BigDiamonds.Rda")
load(rawConnection(diamondsurl))
```

The code used to obtain the data is available here:
https://github.com/solomonm/diamonds-data

## Building a Model Using the Big Diamonds Data Set
Notes:

```{r Building a Model Using the Big Diamonds Data Set}

```


***

## Predictions

Example Diamond from BlueNile:
Round 1.00 Very Good I VS1 $5,601

```{r}
#Be sure you’ve loaded the library memisc and have m5 saved as an object in your workspace.
thisDiamond = data.frame(carat = 1.00, cut = "V.Good",
                         color = "I", clarity="VS1")
modelEstimate = predict(m5, newdata = thisDiamond,
                        interval="prediction", level = .95)
```

Evaluate how well the model predicts the BlueNile diamond's price. Think about the fitted point estimate as well as the 95% CI.

***

## Final Thoughts
Notes:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

